---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# netidmtpreg

<!-- badges: start -->
[![Lifecycle Status](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://lifecycle.r-lib.org/articles/stages.html)
[![Tests](https://github.com/qmarcou/netidmtpreg/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/qmarcou/netidmtpreg/actions/workflows/test-coverage.yaml)
[![R-CMD-check](https://github.com/qmarcou/netidmtpreg/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/qmarcou/netidmtpreg/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of netidmtpreg is to enable net survival estimation through direct
binomial regression, allowing modeling continuous covariate effects that could
not be handled through stratified Pohar-Perme estimation, all this in a 
multistate Illness-Death setting.   

## Installation

You can install the development version of netidmtpreg like so:

``` r
devtools::install_github("qmarcou/netidmtpreg")
```

If working from source through a git clone, the package should be installed locally to get tests running `future::multisession` planning working.
This can be accomplished using `devtools`
```{r eval=FALSE}
devtools::dev_mode(on = TRUE)
devtools::install_local(force = TRUE) # force package update
devtools::load_all() # required to make tests visible
testthat::test_package("netidmtpreg")
# or testthat::test_check("netidmtpreg") to run R CMD check
```

## Example

This is a basic example illustrating net survival estimation on data simulated 
using the package. Let's first generate some data:

```{r Generate random data}
library(netidmtpreg)
library(tidyverse)

n_ind <- 1e2 # number of simulated individuals

# Generate exponentially distributed event times without censoring
synth_idm_data <- generate_uncensored_ind_exp_idm_data(
  n_individuals = n_ind,
  lambda_illness = 1.0,
  lambda_death = 0.1
)

# Generate random age and sex labels
synth_idm_data <-
  synth_idm_data %>% tibble::add_column(
    sex = ifelse(rbinom(n_ind, 1, prob = .5), "male", "female"),
    age = runif(n = n_ind, min = 50, max = 80)
  )

# Generate random start of follow up dates
synth_idm_data <-
  synth_idm_data %>% tibble::add_column(start_date = as.Date.numeric(
    x = rnorm(n = n_ind, mean = 0, sd = 1e2),
    origin = as.Date("15/06/1976", "%d/%m/%Y")
  ))

# Generate population mortality assuming equal constant population rate
l_pop_death <- 0.1 # extra disease mortality doubles the population mortality
population_death_times <- generate_exponential_time_to_event(
  n_individuals = n_ind,
  lambda = l_pop_death
)
# create a corresponding ratetable object
const_ratetable <- survival::survexp.us
const_ratetable[] <- l_pop_death # ratetable's covariate do not matter
# Update death time accordingly to create an observed crude survival dataset
crude_synth_idm_data <- netidmtpreg:::apply_iddata_death(
  synth_idm_data,
  population_death_times
)
```

Now let's carry the of net survival:
```{r estimation}
# Estimation can be sped up and carried in parrallel using futures:
future::plan("multisession") # will work on any OS
# future::plan("multicore") # more efficient but only works on UNIX systems
net_estimate <- renewnetTPreg(
  formula = ~ sex,  # intercept + binary covariate
  data = crude_synth_idm_data,
  # Use a standard ratetable
  ratetable = const_ratetable,
  rmap = list(
    age = age,
    sex = sex,
    year = start_date
  ),
  time_dep_popvars = list("age", "year"),
  s = 0.5,
  trans = "13",
  link = "logit",
  R = 100 # Number of bootstraps
)
future::plan("sequential") # close the multisession, see future's documentation
```

We obtain a `TPreg` object with a dedicated plotting method using ggplot:

```{r plot, echo = FALSE}
plot(net_estimate)
```

More examples to come!